Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2019-06-12 papers/lime_diagnostics/visual-diagnostics-of-a-model-explainer.Rnw
updated readme and paper folder names

\usepackage{xcolor}
\definecolor{blue}{rgb}{0.22, 0.42, 0.73}
\definecolor{orange}{rgb}{0.94, 0.59, 0.19}
\newcommand{\hh}[1]{\textcolor{orange}{#1}}
\newcommand{\kg}[1]{\textcolor{blue}{#1}}

@@

\hh{XXX make the intro a bit more gentle - how about something along the lines that in statistics there are two main uses for models: interpretation and prediction. In machine learning the focus of models is often on prediction ... while these models are highly predictive, this comes at the cost of interpretability.}
Machine learning models have been proven to provide accurate predictions in a wide range of problems. This impressive predictive performance, however, comes at a cost. These models are typically complex and lack a functional form, so the ability to directly interpret the models is lost. This has led machine learning models be referred to as black box models. In many areas of application, the need to be able to explain the results from a model is critical. As a result, an area of research on how to explain the output from machine learning algorithms has emerged.

The forensics sciences and the heath care industry are two areas that have benefitted from machine learning but have an obvious need for explainable predictions. For example, \citet{aoas} discuss the use of a random forest model in the forensics sciences to determine whether two bullets were fired from the same gun. \citet{yu:2018} describe how healthcare has advanced through the use of machine learning in various ways such as using neural networks to automate medical image diagnoses and implementing Bayesian networks to predict clinical outcomes. In \hh{both} of these cases, the decisions made based on the results of the machine learning models can greatly impact the lives of human beings. If it is not possible to explain the output from a model, it becomes questionable whether the results from the model can be trusted to make important decisions.

As an example, suppose that forensics examiners are trying to determine if a bullet used in a crime was fired from the gun that belongs to a suspect in the investigaion. A random forest model could be used to produce a probability of the bullet being fired from the gun under question. Even if the model may return a high probability of a match, it may be difficult for a jury to trust the results from the model when deciding whether to convict the suspect without the forensics examiners having the ability to explain which factors in the model led to this high probability.

@@

One approach to remedy the problem of uninterpretable models is to develop a model explainer that can provide insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model and output from the model to provide understanding into mechanism of how the model produces predictions. Recently, a handful of model explainers have been developed (reference). LIME (local interpretable model explanations) is one such model explainer (reference).

@@

The explanation that is produced from the explainer is also a prediction or an element of a model. Since it is not directly extracted from the complex model, it is important to ask the question, ``How reliable is the explanation?". Currently, there are no recommendations for how to assess the explanations from LIME (need to check to make sure this is a true statement). In this paper, we will present some visualizations tools that can be used to assess the explanations from LIME in classification problems. Additionally, we will show an example using these tools on explanations from a forensics bullet matching dataset, and we will discuss how the results from the example provide some possible insights into the workings of LIME.
\hh{still to come: outline of the remainder of the paper.}

@@

\subsection{Visualization Diagnostic Tools for LIME}

\section{Application of LIME to Bullet Matching Data}

\section{Results}

@@

\subsection{LIME Package Explanations Dependent on Sampling Method}

>>>>

\usepackage{xcolor}
\definecolor{purple}{rgb}{0.5, 0.0, 0.5}
\definecolor{orange}{rgb}{0.94, 0.59, 0.19}
\newcommand{\hh}[1]{\textcolor{orange}{#1}}
\newcommand{\kg}[1]{\textcolor{purple}{#1}}

@@

\hh{In the field of statistics, there are two main uses for models: interpretation and prediction.} \kg{Machine learning models are often used for the latter purpose. While these models have been proven to provide accurate predictions in a wide range of problems, this predictive ability comes at the cost of interpretability.} These models are typically complex and lack a functional form, so the ability to directly interpret the models is lost. \kg{This has earned machine learning models the reputation as being black boxes. The ability to interpret a model helps to diagnose the model and understand the predictions.} In some areas of application, the need to be able to explain the results from a model is critical. \kg{(These last few sentences could use some work so that the ideas are better connected from one sentence to the next.)} As a result, an area of research on how to explain the output from machine learning algorithms has emerged.

The forensics sciences and the heath care industry are two areas that have benefitted from machine learning but have an obvious need for explainable predictions. For example, \citet{aoas} discuss the use of a random forest model in the forensics sciences to determine whether two bullets were fired from the same gun. \citet{yu:2018} describe how healthcare has advanced through the use of machine learning in ways such as using neural networks to automate medical image diagnoses and implementing Bayesian networks to predict clinical outcomes. In \hh{both} of these cases, the decisions made based on the results of the machine learning models would greatly impact the lives of human beings. If it is not possible to explain the output from a model, it becomes questionable whether to trust the results from the model to make decisions.

As an example, suppose that forensics examiners are trying to determine if a bullet used in a crime was fired from the gun that belongs to a suspect in the investigaion. A random forest model could be used to produce a probability that the bullet was fired from the gun under question. Even if the model returns a high probability of a match, it will be difficult for a jury to trust the results from the model when deciding whether to convict the suspect without the forensics examiners having the ability to explain which factors in the model led to the high probability.

@@

One approach to remedy the problem of uninterpretable models is to develop a model explainer that provides insight into the performance of the complex model. A model explainer is a method that is separate from the model but uses the model and output from the model to provide understanding into mechanism of how the model produces predictions. Recently, a handful of model explainers have been developed (reference). LIME (local interpretable model explanations) is one such model explainer (reference).

@@

The explanation that is produced from the explainer is also a prediction or an element of a model. Since it is not directly extracted from the complex model, it is important to ask the question, ``How reliable is the explanation?". Currently, there are no recommendations for how to assess the explanations from LIME (need to check to make sure this is a true statement). \kg{I'd like to expand on this paragraph since this is really the focus of the paper. Then the last paragraph in the intro can outline the paper.}

In this paper, we will present some visualizations tools to assess the explanations from LIME in classification problems. Additionally, we will show an example using these tools on explanations from a forensics bullet matching dataset, and we will discuss how the results from the example provide some possible insights into the workings of LIME.
\hh{still to come: outline of the remainder of the paper.}

\section{Background}

@@

\subsection{Diagnostic Tool 1}

\subsection{Diagnostic Tool 2}

\subsection{etc...}

\section{Application}

@@
